---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# poems: Pattern-oriented ensemble modeling system (for spatially explicit populations)

<!-- badges: start -->
<!-- badges: end -->

The poems package provides a framework of interoperable *R6* (Chang, 2020) classes
for building ensembles of viable models via the pattern-oriented modeling (POM)
approach (Grimm et al., 2005). The package includes classes for encapsulating and 
generating model parameters, and managing the POM workflow. The workflow includes:

1. Model setup including generated spatial layers and demographic population 
   model parameters.
1. Generating model parameters via Latin hypercube sampling (Iman & Conover, 1980).
1. Running multiple sampled model simulations.
1. Collating summary results metrics via user-defined functions.
1. Validating and selecting an ensemble of models that best match known patterns.

By default, model validation and selection utilizes an approximate Bayesian computation 
(ABC) approach (Beaumont et al., 2002) using the *abc* package (Csillery et al., 2015). 
However, alternative user-defined functionality could be employed. 

The package includes a spatially explicit demographic population model simulation
engine, which incorporates default functionality for density dependence, correlated 
environmental stochasticity, stage-based transitions, and distance-based dispersal. 
The user may customize the simulator by defining functionality for trans-locations, 
harvesting, mortality, and other processes, as well as defining the sequence order 
for the simulator processes. The framework could also be adapted for use with other 
model simulators by utilizing its extendable (inheritable) base classes.

## Installation

You can install the released version of poems from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("poems")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("GlobalEcologyLab/poems")
```
## Example

The following simple example demonstrates how to run a single spatially explicit 
demographic population model using *poems*:

```{r example, message = FALSE, fig.align = "center"}
library(poems)

# Demonstration example region (U Island) and initial abundance
coordinates <- data.frame(x = rep(seq(177.01, 177.05, 0.01), 5),
                          y = rep(seq(-18.01, -18.05, -0.01), each = 5))
template_raster <- Region$new(coordinates = coordinates)$region_raster # full extent
template_raster[][-c(7, 9, 12, 14, 17:19)] <- NA # make U Island
region <- Region$new(template_raster = template_raster)
initial_abundance <- seq(0, 300, 50)
raster::plot(region$raster_from_values(initial_abundance), 
             main = "Initial abundance", xlab = "Longitude (degrees)", 
             ylab = "Latitude (degrees)", zlim = c(0, 300), colNA = "blue")

# Set population model
pop_model <- PopulationModel$new(
  region = region,
  time_steps = 5,
  populations = 7,
  initial_abundance = initial_abundance,
  stage_matrix = matrix(c(0, 2.5, # Leslie/Lefkovitch matrix
                          0.8, 0.5), nrow = 2, ncol = 2, byrow = TRUE),
  carrying_capacity = rep(200, 7),
  density_dependence = "logistic",
  dispersal = (!diag(nrow = 7, ncol = 7))*0.05,
  result_stages = c(1, 2))

# Run single simulation
results <- population_simulator(pop_model)
results # examine
raster::plot(region$raster_from_values(results$abundance[,5]),
             main = "Final abundance", xlab = "Longitude (degrees)", 
             ylab = "Latitude (degrees)", zlim = c(0, 300), colNA = "blue")
```

Further examples utilizing the POM workflow and more advanced features of *poems*
can be found in the accompanying vignettes.

## References

Beaumont, M. A., Zhang, W., & Balding, D. J. (2002). 'Approximate Bayesian computation
in population genetics'. *Genetics*, vol. 162, no. 4, pp, 2025–2035.
<doi:10.1093/genetics/162.4.2025>

Chang, W. (2020). 'R6: Encapsulated Classes with Reference Semantics'. *R package version 2.5.0*. 
Retrieved from  <https://CRAN.R-project.org/package=R6>

Csillery, K., Lemaire L., Francois O., & Blum M. (2015). 'abc: Tools for Approximate 
Bayesian Computation (ABC)'. *R package version 2.1*. Retrieved from 
<https://CRAN.R-project.org/package=abc>

Grimm, V., Revilla, E., Berger, U., Jeltsch, F., Mooij, W. M., Railsback, S. F.,
Thulke, H. H., Weiner, J., Wiegand, T., DeAngelis, D. L., (2005). 'Pattern-Oriented
Modeling of Agent-Based Complex Systems: Lessons from Ecology'. *Science*
vol. 310, no. 5750, pp. 987–991. <doi:10.1126/science.1116681>

Iman R. L., Conover W. J. (1980). 'Small sample sensitivity analysis techniques for computer 
models, with an application to risk assessment'. *Commun Stat Theor Methods* A9, pp. 1749–1842.
<doi:10.1080/03610928008827996>

