---
title: "Population introduction via translocation functions"
author: "Global ChEC Lab"
output: rmarkdown::pdf_document
geometry: margin=2cm
vignette: >
  %\VignetteIndexEntry{Translocation Example}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  fig.show = "asis",
  fig.align = "centre",
  results = "markup",
  comment = "#>"
)
```

In this vignette we present a simple example of the *poems* workflow to
show how to use a custom `translocation` function to introduce
populations at defined time steps. The vignette also shows how to define
different growth rates for populations in different regions.

Note that the examples are for a single stage matrix model.

## Setup

As before, We begin by loading the necessary packages and setting a
temporary output directory.

```{r setup}
library(poems)
library(raster)
library(sf)
library(scales)
library(stringi) # for randomly generating file names.

OUTPUT_DIR <- tempdir()

# function to round to any arbitrary value
round_any <- function(x, accuracy, f = round) {
  f(x/ accuracy) * accuracy
}
```

\newpage

## Step 1: Build the population model for the study region

Create a model template using the *PopulationModel* class. This model is
spatially explicit, so is defined using the *Region* class. In addition
we make the model **temporally** explicit so we can introduce
populations at defined time steps.

### Study region

First, we'll define our study region. For this example, we utilise a
`raster::RasterLayer()` of Thylacine habitat suitability.

```{r, message = FALSE, warning=FALSE, results='markup', fig.show='asis',fig.align = "center", fig.width = 4, fig.height = 4}
# Region raster
tasmania_raster <- poems::tasmania_raster
tasmania_raster

# Equal area projection
tasPrj <- 'PROJCS["Tasmania_Lambert_Azimuthal",
                 GEOGCS["GCS_WGS_1984",
                        DATUM["D_WGS_1984",
                              SPHEROID["WGS_1984",6378137.0,298.257223563]],
                        PRIMEM["Greenwich",0.0],
                        UNIT["Degree",0.0174532925199433]],
                 PROJECTION["Lambert_Azimuthal_Equal_Area"],
                 PARAMETER["False_Easting",0.0],
                 PARAMETER["False_Northing",0.0],
                 PARAMETER["Central_Meridian",147],
                 PARAMETER["Latitude_Of_Origin",-42.2],
                 UNIT["Meter",1.0]]'

# Template raster to project to
tempExt <- projectExtent(tasmania_raster, tasPrj)
res(tempExt) <- 10000 # 10 km resolution
tempExt

# Project the region
tasmania_raster <- projectRaster(tasmania_raster, tempExt, 
                                 method = "ngb")
plot(tasmania_raster, main = "Tasmania raster", 
     legend = FALSE,
     col = "#2E8B57", colNA = "grey75")
```


\newpage

Now we can define our `poems::Region`

```{r}
# Tasmania study region (735 non-NA cells stored in the order shown) #
region <- Region$new(template_raster = tasmania_raster)
region$region_raster

# Establish HS template and starting location #
# This will be our initial introduction point
int_ll <- sf_project(from = "EPSG:4326",
                     to = tasPrj,
                     pts = cbind(146.44, -41.18))
int_point <- region$region_indices[
  which(region$region_indices == 
          cellFromXY(tasmania_raster, xy = int_ll))]

# row which corresponds to initial introduction site
int_index <- which(region$region_indices == int_point) # 114

# plot of region, and introduction locations
plot(region$region_raster, 
     main = "Tasmanian study region (cell indices)",
     colNA = "grey75",
     addfun = function() {
       points(xyFromCell(region$region_raster, int_point), col = "red", pch = 16)
     })
```

### Land-use modifier

Here we read in a land-use modifier layer which we can use to make our
region spatiotemporally explicit. This has the effect of altering the HS
values through time which causes dynamic changes in habitat suitability,
and thus population abundances.

While we could change these values now and supply them as is to the
simulator. In this example, we will use a *Generator* object later on to
apply the HS scaling.

```{r}
# read in the land-use modifier
hs_mod <- poems::tasmania_modifier

plot(hs_mod, zlim = c(0, 1), colNA = "grey75",
     col = hcl.colors(100, "RdYlGn"))

# Habitat suitability
hs_raster <- poems::thylacine_hs_raster
hs_raster <- projectRaster(hs_raster, region$region_raster, method = "bilinear")
hs_raster <- stretch(hs_raster, minv = 0, maxv = 1)
hs_raster

# initial_hs needed for generator
initial_hs <- hs_raster <- stack(replicate(n = nlayers(hs_mod), hs_raster))
```

\newpage

While we have read in the land-use modifier, we have not applied it to
the HS values yet. This means that our HS values are still static as
shown by the plot below.

```{r}
# static HS for the moment
## capacity generator will make it temporally dynamic
plot(hs_raster, zlim = c(0, 1), colNA = "grey75",
     col = hcl.colors(100, "RdYlGn"),
     addfun = function() {
       points(xyFromCell(region$region_raster, int_point), pch = 16, cex = 0.5)
     })
```

\newpage

### Environmental correlation

Next, we'll define a distance-based spatial correlation for applying
environmental stochasticity within our model. The generated correlation
data is compacted for computational efficiency (with large-scale
models).

```{r, message = FALSE}
# Distance-based environmental correlation (via a compacted Cholesky decomposition)
env_corr <- SpatialCorrelation$new(region = region, 
                                   amplitude = 0.496, 
                                   breadth = 80, 
                                   distance_scale = 1000)
env_corr$calculate_compact_decomposition(decimals = 4)
```

### Spatially-varying growth rates

Here we use IBRA regions to define varying growth rates for different
populations. These could be expected to occur for example in a
wide-ranging species that occurs in varying habitats (e.g. red foxes in
Australia).

```{r}
# allow growth rates to vary by region using IBRA regions
# Tasmania study Interim Bioregionalisation of Australia (IBRA) bioregion cell distribution
ibra_raster <- poems::tasmania_ibra_raster
ibra_raster <- projectRaster(ibra_raster, region$region_raster, method = "ngb")
plot(ibra_raster, colNA = "grey75",
     breaks = seq(1, 9, 1),
     main = "IBRA regions of Tasmania",
     col = hcl.colors(10, "Lajolla"))

ibra_data <- poems::tasmania_ibra_data
ibra_data

# Calculate cell indices and counts for IBRA bioregions
ibra_indices <- lapply(as.list(ibra_data$index),
                       function(i) {
                         which(ibra_raster[region$region_indices] == i)
                       })
str(ibra_indices)

ibra_polygons <- rasterToPolygons(ibra_raster, dissolve = TRUE, na.rm = TRUE)
ibra_polygons@data <- merge(ibra_polygons@data, ibra_data, 
                            by.x = "layer", by.y = "index")
ibra_polygons

plot(ibra_polygons, col = hcl.colors(9, "Lajolla"), border = "black")
text(ibra_polygons, labels = "abbr", cex = 1.2, halo = TRUE)

rmax_regional <- ibra_raster

# seed is set to keep example results constant
{set.seed(27); rmax <- round(rlnorm(9, 0.94, 0.3), 1)}
for (val in 1:9) {
  rmax_regional[rmax_regional == val] <- rmax[val]
}
plot(rmax_regional, colNA = "grey75",
     legend = FALSE, main = "regional growth rates",
     zlim = range(rmax),
     addfun = function() {
       plot(ibra_polygons, border = hcl.colors(9, "Lajolla"), 
            col = NA, add = TRUE)
       text(ibra_polygons, labels = rmax, halo = TRUE)
     },
     col = hcl.colors(100, "Zissou"))

# set upper and lower growth rates per region
ibra_rmax <- cbind(ibra_data, rmax_lower = round(rmax * 0.6, 2),
                   rmax_mean = round(rmax, 2),
                   rmax_upper = round(rmax / 0.75, 2))
ibra_rmax
```

\newpage

## Step 2: Setup the translocation function

Here we generate a custom *Translocation* class object. This object can
be used to translocate populations from one location to another, or as
shown here, to introduce populations at defined locations and timesteps.
The function could also be expanded to introduce varying number of
animals at each site. For simplicity sake, this example releases the
same number of animals at all locations.

### Introduction sites and times

Here we define the introduction sites by matching
`region$region_indices` to our introduction locations.

We also define the timesteps that the introductions should occur. These
timesteps are sequential from 1, and are not defined by, for example,
calendar years.

```{r, echo=TRUE}
# set up translocation locations and order
intro_trans_ll <- sf_project(from = "EPSG:4326",
                       to = tasPrj,
                       pts = cbind(c(148.01, 144.7, 147.9, 148.27, 145.24), 
                                   c(-40.8, -40.7, -43.2, -42.02, -42.3)))
intro_trans_ll
intro_trans_point <- region$region_indices[which(region$region_indices %in% 
                                             cellFromXY(region$region_raster, 
                                                        xy = intro_trans_ll))]
intro_trans_point <- intro_trans_point[-1]
intro_cells <- intro_trans_point; intro_cells

intro_times <- c(2, 3, 6, 8)

# Introduction times and locations
cbind(intro_times, intro_cells)

plot(region$region_raster,
     main = "Introduction sites",
     col = hcl.colors(100, "Lajolla"),
     addfun = function() {
       plot(ibra_polygons, border = "black", col = NA, add = TRUE)
       points(xyFromCell(region$region_raster, intro_cells), pch = 16, 
              cex = 1.5, col = c("darkgreen", "blue2", "black", "goldenrod"))
       points(region$coordinates[which(region$region_indices %in% intro_cells), ],
       col = "firebrick", cex = 1.5, lwd = 2)
       })
```

\newpage

### Translocation function

Here we define the custom `translocation` function. It's simply a `list`
object, with a nested function that uses parameters from the model to
make changes to the simulated populations.

```{r}
# User-defined translocation function (list-nested) and alias ####

translocation <- list(
  
  # Function parameters (passed to function in params list)
  intro_cells = intro_cells, # cells where pops are introduced
  intro_timesteps = intro_times, # timesteps when introduced
  trans_n = 50, # translocated abundances. If not provided by LHS == 50
  region_indices = region$region_indices,
  
  # Function definition
  translocation_function = function(params) {
    # Unpack parameters (used at every time step)
    intro_cells <- params$intro_cells
    intro_timesteps <- params$intro_timesteps
    simulator <- params$simulator
    stages <- params$stages
    populations <- params$populations
    abundances <- params$abundance
    region_indices <- params$region_indices
    tm <- params$tm # timestep
    sa <- params$stage_abundance
    trans_n <- params$trans_n
    # if introduction at timestep, introduce pops
    if (tm %in% intro_timesteps) {
      # take stage abundance at timestep
      new_sa <- array(sa, c(stages, populations))
      # identifies location of introduction
      trans_loc <- which(region_indices == intro_cells[which(intro_timesteps == tm)])
      # add n individuals regardless of K
      new_sa[trans_loc] <-  new_sa[trans_loc] + trans_n
      return(new_sa)
    } else {
      # else return pops as they are
      new_sa <- array(sa, c(stages, populations))
      return(new_sa)
    }
  }
)
translocation_aliases <- list(intro_cells = "translocation$intro_cells",
                              intro_times = "translocation$intro_timesteps",
                              trans_n = "translocation$trans_n",
                              region_indices = "translocation$region_indices")
```

\newpage

## Step 3: Build generators for dynamically generating model parameters

### Growth rate generator

Now we can build a *Generator* class object that will generate random
growth rates based on quantiles between the min and max in each IBRA
region. This allows us to sample across the range of values within each
region.

```{r}
# Build a Rmax generator based on sampled IBRA Rmax range quantile
rmax_gen <- Generator$new(
  description = "Rmax",
  spatial_correlation = env_corr,
  generate_rasters = FALSE,
  ibra_data_rmax = ibra_rmax,
  ibra_indices = ibra_indices,
  region_cells = region$region_cells,
  inputs = c("rmax_quantile"),
  outputs = c("growth_rate_max"),
  generative_requirements = list(growth_rate_max = "function"))

# growth_rate_max template
rmax_gen$add_function_template(
  "growth_rate_max",
  function_def = function (params) {
    growth_rate_max <- array(0, params$region_cells)
    for (i in 1:nrow(params$ibra_data_rmax)) {
      growth_rate_max[params$ibra_indices[[i]]] <- 
        stats::qunif(params$rmax_quantile, 
                     min = params$ibra_data_rmax$rmax_lower[i],
                     max = params$ibra_data_rmax$rmax_upper[i])
    }
    return(growth_rate_max)
  },
  call_params = c("ibra_data_rmax", "ibra_indices", "region_cells", "rmax_quantile"))

# test rmax generator at median values
rmax_gen_ex <- rmax_gen$generate(input_values = list(rmax_quantile = 0.5))
rmax_regional[region$region_indices] <- rmax_gen_ex$growth_rate_max
plot(rmax_regional, main = "median regional rmax",
     col = hcl.colors(100),
     addfun = function() {
       plot(ibra_polygons, border = "black", col = NA, add = TRUE)
     })
```

The map above shows the result of sampling our possible *rmax* values to
their median values. The plot below shows that we can sample across a
range of values using the same generator. This functionality allows us
to pass `rmax_quantile` as a variable in a *LatinHypercubeSampler*
class.

```{r}
# Test multiple quantiles
test_rmax <- lapply(seq(0, 1, 0.1), function(i) {
  region$raster_from_values(rmax_gen$generate(input_values = list(rmax_quantile = i))$growth_rate_max)
})
test_rmax <- stack(test_rmax)
names(test_rmax) <- paste0("Q", seq(0,1,0.1))

# plot
plot(test_rmax, colNA = "grey75",
     legend = TRUE,
     zlim = c(min(values(test_rmax), na.rm = TRUE),
              max(values(test_rmax), na.rm = TRUE)),
     addfun = function() {
       plot(ibra_polygons, border = "black", col = NA, add = TRUE)
     },
     col = hcl.colors(100))
```

\newpage

### Dispersal generator

This generator controls dispersal across our region based on mean
dispersal distance and proportion of dispersers.

```{r}
# Dispersal generator ####
# Set for veriable mean distance, max hard-coded at 150
dispersal_gen <- DispersalGenerator$new(
  region = region,
  spatial_correlation = env_corr,
  generate_rasters = FALSE,
  dispersal_max_distance = 150,
  distance_classes = seq(5, 150, by = 10),
  distance_scale = 1000, # in km
  dispersal_friction = DispersalFriction$new(), # modify coastline distances
  inputs = c("dispersal_p", "dispersal_b"), # proportion and average distance
  decimals = 4)
dispersal_gen$calculate_distance_data()
head(dispersal_gen$distance_data$base, 10)
table(dispersal_gen$distance_data$base$distance_class)
```

Like with the `rmax_gen` defined above, we can test our dispersal
generator over a range of values to make sure the dispersal curve looks
appropriate. We can see in the plot below that as the proportion of
dispersal decreases, so does the maximum dispersal distance.

```{r}
# plot dispersal curves for mean dispersal rates
disp_fun <- function(p, b, distance) {
  p * exp(-distance/b)
}

disp_mat <- data.frame(p = round(runif(1000, 5, 40)/100,2), # prop
                       b = round(runif(1000, 5, 40)) # mean distance
                       )
head(disp_mat)

disp_test <- lapply(1:nrow(disp_mat), function(i) {
  p <- disp_mat[i, "p"]
  b <- disp_mat[i, "b"]
  disp_x <- disp_fun(p, b, seq(5, 150, 5))
  return(disp_x)
})

{par(mar = c(4,4,0.5,0.5))
  matplot(x = seq(5, 150, 5), y = rep(NA, 30), type = "l", ylim = c(0, 0.4),
        xlab = "Disp. dist (km)", ylab = "Prop. disp.", yaxt = "n", xaxt = "n")
  axis(1, at = seq(0, 150, 10))
  axis(2, at = seq(0, 40, 5)/100, labels = seq(0, 40, 5))
  lapply(disp_test, function(i) {
    matplot(x = seq(5, 150, 5), y = unlist(i), type = "l", add = TRUE,
            col = c("#C9C9C944"))}
    )
  lines(x = seq(5, 150, 5),
        y = apply(as.data.frame(disp_test), 1, mean), col = "firebrick")
  }
dev.off()
```

Now that we're happy the curve follows the right form, we can generate
some dispersal data to test our population model with.

```{r}
# Generate sampled dispersals for p = 0.35, b = 40 (km)
sample_dispersal_data <- dispersal_gen$generate(
  input_values = list(dispersal_p = 0.35, dispersal_b = 40))$dispersal_data
head(sample_dispersal_data[[1]], 10) # examine
```

\newpage

### Capacity generator

Here we build a carrying capacity generator. Carrying capacity is based
on maximum density values and scaled by HS (i.e. cells with a HS of 1
contain the highest densities).

The capacity generator is set-up in such a way that it requires multiple
parameters to work.

1.`max_dens`: the maximum theoretical density of populations

2.`q_thresh`: the quantile threshold used to rescale the HS values

3.`trans_n`: the number of animals that are introduced. This value is
consistent with the `translocation` function.

```{r}
capacity_gen <- Generator$new(
  description = "capacity",
  spatial_correlation = env_corr,
  generate_rasters = FALSE,
  time_steps = ncol(initial_hs),
  hs_raster = initial_hs[region$region_indices], # provide full stack of HS. Template attached
  hs_mod = hs_mod[region$region_indices], # provide full stack of LULC modifier. Template attached
  int_index = int_point,
  trans_n = translocation$trans_n, # number of animals introduced
  region_indices = region$region_indices,
  inputs = c("max_dens", "q_thresh", "trans_n"),
  outputs = c("initial_abundance", "carrying_capacity"),
  generative_requirements = list(initial_abundance = "function",
                                 carrying_capacity = "function"))

capacity_gen$add_function_template(
  param = "initial_abundance",
  function_def = function (params) {
    distr_a <- params$hs_raster[, 1]
    ## 0 everywhere except the intro point at the first time step
    ## intro point to trans_n
    ## Could be above or below carrying capacity
    idx <- which(params$region_indices == params$int_index)
    distr_a[idx] <- params$trans_n
    distr_a[-idx] <- 0
    return(distr_a)
  },
  call_params = c("hs_raster", "int_index", "region_indices", "trans_n"))

capacity_gen$add_function_template(
  "carrying_capacity",
  function_def = function (params) {
    idx <- which(params$region_indices == params$int_index)
    distr_k <- params$hs_raster
    distr_mod <- params$hs_mod
    stopifnot("hs_raster and hs_mod have different number of layers" =
                dim(distr_k) == dim(distr_mod))
    # stretch HS values based on q_thresh
    distr_k <- scales::rescale(distr_k, from = c(0, params$q_thresh), to = c(0, 1))
    distr_k[distr_k < 0] <- 0
    distr_k[distr_k > 1] <- 1
    # multiply thresholded HS by hs_modifier
    distr_k <- distr_k * distr_mod
    # rescale back to {0, 1}
    qMax <- max(distr_k, na.rm = TRUE)
    distr_k <- scales::rescale(distr_k, from = c(0, qMax), to = c(0, 1))
    distr_k[distr_k < 0] <- 0
    distr_k[distr_k > 1] <- 1
    # carrying capacity = (HS * maximum density)
    distr_k <- ceiling(distr_k * params$max_dens)
    distr_k[idx, 1] <- params$max_dens
    # distr_k[-idx, 1] <- 0
    return(distr_k)
  },
  call_params = c("hs_raster", "hs_mod", "int_index", "region_indices", "max_dens", "q_thresh"))

# have all parameters been specified correctly
capacity_gen$generative_requirements_satisfied()
```

Now we have defined our generator we can make some test data for the
model

```{r}
# Generate example initial abundance and declining carrying capacity time-series
generated_k <- capacity_gen$generate(input_values = list(max_dens = 100, q_thresh = 0.90, 
                                                         trans_n = 60))
example_initial_abundance <- generated_k$initial_abundance
example_carrying_capacity <- generated_k$carrying_capacity

# Plot the example initial abundance
example_initial_n_raster <- region$raster_from_values(example_initial_abundance)
example_initial_n_raster
plot(example_initial_n_raster, main = "Example initial abundance",
     col = hcl.colors(100, "Lajolla", rev = TRUE), colNA = "grey75",
     addfun = function() {
       plot(ibra_polygons, border = "black", col = NA, add = TRUE)
     })

# Plot the carrying capacity
## carrying capacity is forced to maximum theoretical value at first time step
example_k <- region$raster_from_values(example_carrying_capacity)
example_k[[c(1, 6, 11)]]
plot(example_k,
     col = hcl.colors(100, "RdYlGn", rev = TRUE), colNA = "grey75",
     addfun = function() {
       plot(ibra_polygons, border = "black", col = NA, add = TRUE)
     },
     zlim = c(0, 100))
```

\newpage

## Step 4: Build a template model

Using the generators we've built we can now test if our simple
population model works as expected.

```{r}
# Template model ####
model_template <- PopulationModel$new(
  region = region,
  time_steps = 11,
  years_per_step = 1,
  stage_matrix = 1, # single-stage
  populations = region$region_cells, # 735
  demographic_stochasticity = TRUE,
  standard_deviation = 0.18,
  density_dependence = "logistic", # Ricker
  harvest = FALSE, # No harvest
  dispersal = dispersal_gen,
  translocation = translocation,
  dispersal_source_n_k = list(threshold = 0.92, cutoff = 0),
  simulation_order = c("translocation", "results", "transition", "dispersal"),
  random_seed = 20230210,
  attribute_aliases = translocation_aliases,
  results_selection = c("abundance"))

model <- model_template$clone()
model$set_attributes(initial_abundance = example_initial_abundance,
                     carrying_capacity = example_carrying_capacity,
                     growth_rate_max =  rmax_gen_ex$growth_rate_max,
                     translocation = translocation,
                     trans_n = 75, # passed through to translocation function
                     dispersal = sample_dispersal_data)
# run poems simulator
results <- population_simulator(model)
results$all$abundance

# timeseries of total abundance
plot(1:11, results$all$abundance, type = "l", 
     xlab = "timestep", ylab = "Total abundance")
```

The template model runs successfully and now we can make some maps of
the dispersal patterns of the populations from their initial
introduction.

```{r}
abund_ras <- region$raster_from_values(results$abundance)
abund_ras[[c(1, 6, 11)]]
abd_max <- round_any(max(values(abund_ras), na.rm = TRUE), 20, f = ceiling)

# plot of abundances. log(x+1) transformed.
plot(log1p(abund_ras),
     col = hcl.colors(100),
     colNA = "grey75",
     addfun = function() {
       plot(ibra_polygons, border = "black", col = NA, add = TRUE)
     },
     zlim = c(0, log1p(abd_max)))
```

To make sure our `translocation` function is working correctly, we can
run the template model again but with the `translocation` function
turned off.

```{r}
model$set_attributes(initial_abundance = example_initial_abundance,
                     carrying_capacity = example_carrying_capacity,
                     growth_rate_max = rmax_gen_ex$growth_rate_max,
                     translocation = NULL,
                     dispersal = sample_dispersal_data)
results_notransn <- population_simulator(model) # run poems simulator
results_notransn$all$abundance
results$all$abundance
abund_ras_notransn <- region$raster_from_values(results_notransn$abundance)
abund_ras_notransn[[c(1, 6, 11)]]
diff_ras <- abund_ras - abund_ras_notransn
diff_ras[[1:2]]

```

We can see from the `diff_ras` above that the difference between
timesteps 2 of the two model runs is 75 individuals, which was the
value passed through to the `translocation` function of the original
model run.

And now we can plot the difference.

```{r}
plotmax <- round_any(max(abs(values(diff_ras)), na.rm = TRUE), 10, ceiling)
plot(diff_ras[[c(1, 6, 11)]],
     zlim = c(-plotmax, plotmax),
     breaks = c(-plotmax, -100, -50, -20, 0, 20, 50, 100, plotmax),
     col = hcl.colors(9, "PuOr"),
     colNA = "grey75",
     addfun = function() {
       plot(ibra_polygons, col = NA, add = TRUE)
     })
```

\newpage

## Step 5: Run multiple simulations

In order to explore the model parameter space to find the best models,
we generate Latin hypercube samples of model and generator parameters to
be simulated, using the *LatinHypercubeSampler* class. This class has
functionality for generating sample parameters via Uniform, Normal,
Lognormal, Beta, and Triangular distributions. For our example we only
generate 10 samples. Typically however, a user would need to generate
thousands to tens of thousands, of samples.

### Define the latin-hypercube for sampling

```{r}
# Latin-hypercube sampler ####
lhs_gen <- LatinHypercubeSampler$new()

# Habitat suitability threshold
lhs_gen$set_uniform_parameter("q_thresh", lower = 0.90, upper = 0.99, decimals = 2) 

# Growth rate
lhs_gen$set_uniform_parameter("rmax_quantile", lower = 0, upper = 1, decimals = 2)
lhs_gen$set_uniform_parameter("standard_deviation", lower = 0.00, upper = 0.70, decimals = 2)

# Dispersal
lhs_gen$set_uniform_parameter("dispersal_p", lower = 0.05, upper = 0.40, decimals = 2)
## mean dispersal between 5 and 40 km
lhs_gen$set_uniform_parameter("dispersal_b", lower = 5, upper = 40, decimals = 0)
lhs_gen$set_uniform_parameter("dispersal_n_k_threshold", lower = 0.7, upper = 1.0, decimals = 2)

# Density max
## Density: animals/km2 needs to be scaled by grid size (10km x 10km)
## e.g. 1/km2 = (1 animal/km2 * (10*10) ) * frac_cell_used
## 1 km2 = 80 per grid cell = (1*(10*10))*0.8 # assuming 80% grid cell used
## Here I have assumed only 80% of cell is suitable. Upper/lower = 1/km - 6.25/km
lhs_gen$set_uniform_parameter("max_dens", lower = 80, upper = 500, decimals = 0)

# Translocation
lhs_gen$set_uniform_parameter("trans_n", lower = 10, upper = 100, decimals = 0)

sample_data <- lhs_gen$generate_samples(number = 10, random_seed = 42)
head(sample_data)

# Make unique row names for saving files
{set.seed(54612)
  sample_data$UniqueID <- paste0(
    stri_rand_strings(nrow(sample_data), 4, "[A-Z]"),
    stri_rand_strings(nrow(sample_data), 4, "[0-9]"))
}
sample_data <- sample_data[, c(9, 1:8)]
sample_data
```

We we can run a simulation for each set (or row) of sampled parameters.
The *SimulationManager* class manages the generation of parameters (via
the generators), the running the model simulations, and writing
simulation results to disk. It also maintains a log of each simulation's
success and any errors or warnings encountered.

```{r}
model <- model_template$clone()
model$set_attributes(params = list("standard_deviation" = NULL,
                                   "dispersal_source_n_k$threshold" = NULL,
                                   "dispersal_source_n_k$cutoff" = 0.00))

# Build the simulation manager
sim_manager <- SimulationManager$new(
  sample_data = sample_data,
  model_template = model,
  # initial_hs = initial_hs,
  generators = list(dispersal_gen, capacity_gen, rmax_gen),
    parallel_cores = 1L,
  results_filename_attributes =
    c(NULL, "UniqueID", "results"),
  results_ext = ".RDS",
  results_dir = OUTPUT_DIR)

# Takes <10 seconds to run 10 example sims on a single core.
system.time({
  run_output <- sim_manager$run()
})
run_output$summary
```

## Step 6: Extract results from simulations

Now that the simulations have run, we can extract the modelled
abundances. We now wish to collate summary results for each of our
simulations via the *ResultsManager* class. This manager loads the
results from each sample simulation into an intermediate
*PopulationResults* class object, which dynamically generates further
results. We need to define functions for calculating summary metrics, as
well as any matrices (one row of values per simulation) that we may be
interested in examining.

We can see from above that the `run_output$summary` shows that there
were a number of simulations that didn't complete successfully:

```{r}
run_output$summary
```

The `run_output$full_log` shows that some simulations produced NA values
when calculating population abundances:

`"Warning: Non-finite stage abundances returned by user-defined translocation function"`

These simulations had a range of parameter values that caused the
simulation to "fall-over". These errors could be fixed by improving the
`translocation` function, or alternatively the user could discard the
simulations as being structurally wrong - i.e. that specific combination
of parameters is simply unsuitable. We're going to treat them as the
latter.

### Extract results

```{r}
# Extract timeseries of abundance from each of the sims
# Load our results (list) into a PopulationResults object
p_results <- PopulationResults$new(results = run_output)
res_manager <- ResultsManager$new(simulation_manager = sim_manager,
                                  simulation_results = p_results,
                                  generators = NULL,
                                  summary_matrices = c("n",
                                                       "distr_pop"),
                                  summary_functions = list(
                                    # total pop abundance
                                    "n" = function(sim_results) {
                                      sim_results$all$abundance
                                    },
                                    # matrix of abundance
                                    ## can be made into raster
                                    "distr_pop" = function(sim_results) {
                                     sim_results$abundance
                                    }
                                  ),
                                  parallel_cores = 1L)
gen_log <- res_manager$generate()
gen_log$summary

# matrix of total population abundances
## each row is a sim, each column a timestep
res_manager$summary_matrix_list$n

# plot
matplot(x = 1:ncol(res_manager$summary_matrix_list$n), 
        y = t(res_manager$summary_matrix_list$n), type = "b",
        lty = 1, xlab = "timestep", ylab = "total abundance")
```

From the results we can see that our `translocation` function was
working correctly, and that our `capacity_gen` was defined correctly, as
the abundances in the first time step are the same as the `trans_n`
values from our latin-hypercube samples.

```{r}
identical(unlist(res_manager$summary_matrix_list$n[, 1]),
          unlist(sample_data$trans_n))
```

Let's assume that samples 2, 3, 4, and 8 are our "best" simulations
after some sort of validation. We have already extracted the abundances
from these simulations, so now with a few more lines of code, we can
generate averages of the simulations as rasters

```{r}
best_sims <- c(2:4, 8)
dim(res_manager$summary_matrix_list$distr_pop[best_sims, ])

best_abund <- matrix(
  nrow = region$region_cells,
  ncol = 11, # 11 timesteps,
  data = round(colMeans(res_manager$summary_matrix_list$distr_pop[best_sims, ])))

best_abund <- region$raster_from_values(best_abund)
best_abund[[c(1, 6, 11)]]

abd_max <- round_any(max(values(best_abund), na.rm = TRUE),
                     accuracy = 100, ceiling)

# plot of log(x+1) abundances
plot(log1p(best_abund),
     col = hcl.colors(100, "Spectral", rev = TRUE),
     colNA = "grey75",
     addfun = function() {
       plot(ibra_polygons, border = "#000000", col = NA, add = TRUE)
       },
     zlim = c(0, log1p(abd_max)))

```
